---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  fig.path = "man/figures/README-"
  )
```


# subsetnc

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/markwh/subsetnc.svg?branch=master)](https://travis-ci.org/markwh/subsetnc)


[![Codecov test coverage](https://codecov.io/gh/markwh/subsetnc/branch/master/graph/badge.svg)](https://codecov.io/gh/markwh/subsetnc?branch=master)
<!-- badges: end -->

A little package for making new netcdfs as subsets of existing netcdfs

## TODO

- Add examples to function documentation
- add travis, covr, badges

## Motivation

[NetCDF](https://www.unidata.ucar.edu/software/netcdf/) is a commonly used format for storing scientific data. Several useful R packages already exist for reading, writing, and manipulating netcdf data, including [ncdf4]() and [tidync](https://github.com/ropensci/tidync). Unfortunately, none of them facilitate simple subsetting of netcdf files to produce new netcdf files. This is the job of **subsetnc**. 


**subsetnc** features:

- familiar dplyr-style filtering using both variables (1-D only) and dimensions
- support for [groups](https://www.unidata.ucar.edu/software/netcdf/docs/group__groups.html)
- somewhat-optimized variable access (though likely room for improvement)

Possible use cases include:

- making example or unit-test netcdfs from larger files (this was the impetus for making the package)
- extracting only the desired sections of a very large dataset before transferring (e.g. from remote storage to local)
- sharing just the important parts of a large dataset with colleagues, etc. 


## Installation

```{r, eval = FALSE}
devtools::install_github("markwh/subsetnc")
```

## Basic usage

A ncdf4 object can be subset by either its dimension(s) or 1-dimensional variables using logical statements similar to `dplyr::filter()`. Any number of substting statemenst can be combined, for example:

```{r, eval = FALSE}
nc <- ncdf4::nc_open(ncfile)

ssnc <- nc %>% 
  nc_subset(var1 %in% 1:5,
            var2 > var1 + 7,
            var3 %in% (median(var3) + (-3:3)),
            dim1 < 100,
            filename = "newnc.nc")
```


The output of `nc_subset()` is an ncdf4 object, which points to a newly created on-disk netcdf file, specified by the `filename` argument. If `filename` is omitted, the new netcdf is written to a `tempfile()` that will be deleted upon R session exit. The result can therefore be used with any of the `ncdf4::` functions. 


### Added variables

The netcdf produced by `nc_subset()` contains additional variables corresponding to the dimension values in the original netcdf. Because netcdf dimensions are often required to be valued `1:length(dim)` (unless they are specified with units, which in my experience is rare), these new variables are required in order to keep track of dimension indices--to facilitate joining to other datasets, for example. These variables have names identical to their corresponding dimensions, with two underscores appended to the end. For example, the variable `group1/dim1__` is created to keep track of the values of the *dimension* `group1/dim1` from the original netcdf. 

## Examples

An example netcdf dataset is included, showing node- and reach-level simulated [SWOT](https://swot.jpl.nasa.gov/) data from the Sacramento River in California. This dataset includes node- and reach-level measurements of river height, width, slope, and other variables. 

Because the example dataset contains [groups](https://www.unidata.ucar.edu/software/netcdf/docs/group__groups.html), its variables and dimensions must be accessed in the `<group>/<variable>` format. Because of the `/`, these variables and dimensions must be enclosed in backticks (\`) when subsetting. 

```{r}
library(ggplot2)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ncdf4)
library(subsetnc)

ncfile <- system.file("extdata", "sacriver.nc", package = "subsetnc", mustWork = TRUE)
orignc <- nc_open(ncfile)

plotvars <- c("latitude", "longitude", "height")
mapdf_orig <- paste0("nodes/", plotvars) %>% 
  lapply(ncvar_get, nc = orignc) %>% 
  lapply(as.vector) %>% 
  setNames(plotvars) %>% 
  as.data.frame()

mapdf_orig %>% 
  ggplot(aes(x = longitude, y = latitude, color = height)) + 
  geom_point(size = 4) +
  coord_map()
```


### Subset by dimension:

```{r}
# Keep every 3rd node along the `nodes/nodes` dimension
ssnc_dim <- nc_subset(orignc, `nodes/nodes` %% 3 == 0)

mapdf_dimss <- paste0("nodes/", plotvars) %>% 
  lapply(ncvar_get, nc = ssnc_dim) %>% 
  lapply(as.vector) %>% 
  setNames(plotvars) %>% 
  as.data.frame()

mapdf_dimss %>% 
  ggplot(aes(x = longitude, y = latitude, color = height)) + 
  geom_point(size = 4) +
  coord_map()

```


### Subset by variable: 

```{r}

ssnc_var <- nc_subset(orignc, `nodes/width` < median(`nodes/width`))

mapdf_varss <- paste0("nodes/", plotvars) %>% 
  lapply(ncvar_get, nc = ssnc_var) %>% 
  lapply(as.vector) %>% 
  setNames(plotvars) %>% 
  as.data.frame()

mapdf_varss %>% 
  ggplot(aes(x = longitude, y = latitude, color = height)) + 
  geom_point(size = 4) +
  coord_map()

```


### Combine subset expressions (variable and dimension): 


```{r}

ssnc_vardim <- nc_subset(orignc, 
                         `nodes/width` < median(`nodes/width`),
                         `nodes/height` > 8,
                         `nodes/nodes` %% 3 == 0)

mapdf_vardimss <- paste0("nodes/", plotvars) %>% 
  lapply(ncvar_get, nc = ssnc_vardim) %>% 
  lapply(as.vector) %>% 
  setNames(plotvars) %>% 
  as.data.frame()

mapdf_vardimss %>% 
  ggplot(aes(x = longitude, y = latitude, color = height)) + 
  geom_point(size = 4) +
  coord_map()

```


